Understanding the Nanoscale Structure of Inverted Hexagonal Phase Lyotropic Liquid Crystal Polymer Membranes

### 2018-10-10 Ben_Manuscripts/structure_paper/Final_Draft.tex
Follow-up edits

  Reverse osmosis (RO) and nanofiltration (NF) are two prevailing membrane
  filtration processes that can be used to separate solutes on the order of 1 nm
  in size and smaller, including 
  %MRSF: this does seem to blend RO and NF together; is this correct, or are there important differences that should be outlined?
  %BJCF: tried to differentiate them
  ions. However, commercially available amorphous RO membranes and porous NF
  membranes lack uniformity, which severely impacts their
  selectivity \cite{van_der_bruggen_review_2003}. Although scalable, their
  fabrication involves the spontaneous assembly of polymers into disordered
  structures that offer separation pathways that are tortuous and polydisperse
  in size \cite{werber_materials_2016}. Tortuosity increases the effective
  length that a solute must travel, while pore size polydispersity limits membrane
  selectivity.

@@

  Systems built with 5 columns around each pore equilibrate to a pore spacing that
  is most consistent with the experimental value of 4.12 nm derived from SAXS
  measurements (Figure~\ref{fig:SAXS}). Ordered basin systems built with 4
  columns per pore equilibrate to an average pore spacing 0.25 nm lower than
  experiment. Ordered basin systems built with 6 columns per pore, have an
  equilibrated pore spacing ca. 0.50 nm higher than experiment. Monomers in
  disordered basin systems built with 6 columns-per-pore agree with experimental
  pore-to-pore distances within error, but stack too far apart. 
  %MRSF: not clear of the disordered ones below are for 5 or 6 columns
  Monomers in 6 column-per-pore disordered sandwiched and disordered parallel displaced
  configurations stack $\sim$~4.87 and 4.94 \AA~apart respectively, which is 
  $\sim$ 1.2~\AA~further apart than suggested by experiment. 5 column-per-pore systems
  stack, at a maximum, 0.9 \AA~further apart than experiment (see Table~\ref{table:correlation_length}).
  The remainder of this discussion will therefore focus on the analysis of 
  systems built with 5 columns per pore.

@@

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{intensity_vs_zsigma.pdf}
  \caption{}\label{fig:intensity_vs_zsigma}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{rpi_xsection_vs_zsigma.pdf}
  \caption{}\label{fig:rpi_xsection_vs_zsigma}
  \end{subfigure}
  %MRSF: yellow not legended?
  \caption{We can decrease the maximum intensity of R-$\pi$ by increasing
  thermal noise in the $z$-direction; however, the $q_r$ cross-section profile does
  not change. (a) The intensity of R-$\pi$ drops precipitously as we increase
  $\sigma_z$. We can decrease the intensity of R-$\pi$, relative to the
  background intensity, relative to systems with simulation levels of disorder (blue)
  by a factor of 3 when we increase $\sigma_z$ by 12\% (green) and by a factor 
  of 15 when we increase $\sigma_z$ by 27\% (red). (b) Although the maximum intensity of R-$\pi$
  decreases with increasing $\sigma_z$, the peak shape remains the same.}\label{fig:znoise}
  \end{figure}

@@

  %MRSF: I had a bit of hard time connecting the thesis of this paragraph to the details provided.
  %BJCF: tried reorganizing a bit
  It therefore seems likely that the discrepancy between the shape and intensity of
  R-$\pi$ generated from our simulations versus experiment is primarily due to
  highly correlated columns in our highly symmetric initial configurations.
  %MRSF2: These next two sentences gets pretty confusing; too many clauses, if, when, since . . . make sentences shorter. 
  If the columns were uncorrelated, $g(z)$ of the ordered sandwiched system 
  would very closely resemble Figure~\ref{fig:z_correlation_sandwich} when we 
  average all $z$-slices of the 3D correlation function since contributions to
  $g(z)$ from head groups in other columns within the same pore, and with head
  groups in different pores would cancel out. However, when we include all 
  $z$-slices in $g(z)$ for our system, rather than restricting it to those 
  $z$-slices within 2.1~\AA of the center of the correlation function 
  (Section~\ref{section:correlation_length}), $g(z)$ shows oscillatory 
  behavior with a correlation length that nearly triples to 12.4 $\pm$ 0.7 \AA.
  (Figure~\ref{fig:z_correlation_fullbox}). The peaks of $g(z)$ are actually 
  reinforced by head groups in neighboring columns which is not expected
  in a system where columns act independently.
  %Since all columns in our initial
  %configuration are started at the same $z$-coordinate, $g(z)$ on average exhibits
  %correlation of head groups with other head groups within the same column, with
  %head groups in other columns within the same pore, and with head groups in different pores.

  %BJC9: Note that the magnitude of the correlation function is much lower than 
  % in figure 10a. That's because many of the z-slices are zero when you include
  % all of them.
  \begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{z_correlation_fullbox.pdf}
  \caption{When we average all $z$-slices of the 3D correlation function, the 
  correlation length of the ordered sandwich configuration nearly triples.}\label{fig:z_correlation_fullbox}
  \end{figure}

  Increasing thermal noise in the $xy$ plane causes the FWHM of the $q_r$
  cross-section of R-$\pi$ to, somewhat counterintuitively, decrease. Using a 
  system with uncorrelated columns, we modified the disorder in both the $r$ and
  $\theta$ directions by multiplying their values by a factor of 0.5 and 2 
  (Figure~\ref{fig:qy_fwhm}). When we cut the noise in half, the FWHM increases
  by 88\%. When we double the noise, the FWHM decreases by 51\%. As we have fit
  the data in Figure~\ref{fig:sim_rsection_fit}, the FWHM of R-$\pi$ agrees
  well with experiment, meaning the quenched configuration contains about the
  right amount of disorder on the $xy$ plane. However, the relatively low quality
  of the fit implies that the FWHM of a fixed Lorentzian fit to data from a 
  single simulation might not be the best comparison to experiment and that 
  some form of initial configuration dependence causes extra structure.
  %BJCF: new ending to this paragraph (above) 
  %may actually be different from experiment, and
  %that the radial ordering of the head groups may benefit from further optimization.

@@

  %BJCF: broke giant paragraph into 3 paragraphs, adjusted theses
  %MRSF2: broke up the sentence, check to make sure it makes sense.
  Since the dynamics of this system are slow, consequently most of our analysis
  has been focused on a set of quenched configurations with some structural 
  characteristics that do not vary significantly with time.  We therefore attempted to learn
  more about their time-averaged structures by studying ensembles of shorter 
  simulations started from more independent initial configurations. To do this, we created
  ensembles of 40 simulations in both the parallel displaced and sandwiched 
  configurations. We used the same equilibration scheme described in Section~\ref{section:equilibration}. 
  However, we built the initial 
  configurations with columns randomly displaced in the $z$-direction and looked at
  the results after 5 ns of unrestrained simulation, which should capture any 
  quenched disorder built in from the initial configurations. We then simulated
  XRD patterns from the 40 trajectories, concatenated together. The $q_r$ 
  cross-sections of R-$\pi$ generated from the two concatenated trajectories are shown
  in Figure~\ref{fig:ensemble_XRD}.

@@

  %MRSF: is this the clearest thesis?  Similar to each other?  Similar to the amount of disordered viewed in the long simulations? Is the symmetric nature more useful?  I think also this paragraph should be split up more to emphasize findings. 
  %BJCF: thesis changed, added reminder of initial randomization.
  These short independent simulations exhibit a symmetric distribution of quenched
  disorder providing evidence that the quenched disorder is randomized in each system. 
  The only difference between each independent initial configuration
  is the random $z$-direction displacement that we apply to each column. The
  quenched disorder is then a result of the initial structural collapse that is 
  directed by the randomized velocity vectors of the system's atoms drawn from
  a Maxwell-Boltzmann distribution. We pooled the deviations of each head group COM 
  from idealized positions from all independent simulations in order to create the 
  distributions shown in Figure~\ref{fig:offset_ensemble_pooled} (see Section~\ref{method:simple_systems}).
  The distributions appear symmetric meaning that there is an equal chance for a head
  group to displace in the positive or negative $z$, $r$, or $\theta$-direction.

  %BJCF: broke into separate paragraph
  We verified that this same symmetry exists for individual COMs.
  If we consider the COM of a given head group, we expect that the distribution of
  its deviations from its idealized position among all the independent simulations
  should be symmetric about its idealized position. We can calculate the 
  empirical cumulative distribution function (ECDF) of these distributions of 
  deviations. The ECDFs of 3 randomly selected COMs are shown in 
  Figure~\ref{fig:offset_ecdfs} to guide the eye (colored lines). The average 
  of all COM ECDFs is represented by the red line with the 95\% confidence
  interval represented by the red error bars. We also calculated 
  the ECDF from the full distributions in Figure~\ref{fig:offset_ensemble_pooled}
  (black line with error bars). We generated error bars that represent the 95\% 
  confidence intervals by performing 1000 trials where we randomly sampled 40
  values (the same as the number of independent simulations) from the distributions in 
  Figure~\ref{fig:offset_ensemble_pooled}, then measured the deviation between
  the sampled ECDF and the full ECDF. These error bars are generally smaller than
  those generated from the 400 COM ECDFs, as there is a wider
  distribution of mean COM deviations than the mean deviations from distributions
  sampled from the full distribution. The similarity between entirely independently
  sampled ECDFs and the ECDFs from individual independent simulations supports 
  the hypothesis that each quenched configuration is random and mostly independent.

@@

  %MRSF: seems like a thesis w/o supporting data - maybe merge paragraphs, or provide a survey of the findings presented below.
  %BJC: merged above

@@

  We observe slow dynamics in our system. Typical diffusion constants for columnar
  liquid crystals have been reported to be between $10^{-11}~ \mathrm{m}^2/\mathrm{s}$ 
  \cite{dong_translational_1984} and $10^{-14}~\mathrm{m}^2/\mathrm{s}$ \cite{dvinskikh_molecular_2002}. 
  %MRSF: see below comment; if not LLC's we should qualify this (electrostatics would likely make slower, since larger
  %energies involved.
  %MRS13: are these hydrophilic/hydrophopic LC?  Ones with electrostatic driving forces are likely 
  % to be even slower.  
  %BJCF: new addtion
  In the case of LLCs, we would expect a diffusion coefficient at the low end of this
  range since hydrophilic regions of monomers can hydrogen bond, thus increasing the energetic 
  barriers for motion. We measured the diffusion constants of monomers in each of the main 
  systems we studied (see Table \ref{S-table:msd} of the SI) and find they are all on the 
  order of $10^{-14}~\mathrm{m}^2/\mathrm{s}$. However, since on this timescale, no monomers 
  move a full `level' above or below, this may be an overestimate. This slow speed is
  not entirely surprising considering how densely the monomers are packed as implied
  by diffraction experiments. The entangled tails likely restrict both vertical and lateral
  movement.

@@

%MRSF: get rid of google books id on ref 45.
  \clearpage
  \bibliography{llc}

@@

  \section*{Graphical TOC Entry}
  \begin{figure}
  \includegraphics[width=4.07cm]{toc_waxs.pdf}
  \includegraphics[width=4.5cm]{toc_image_final.pdf}
  \end{figure}

>>>>

  Reverse osmosis (RO) and nanofiltration (NF) are two prevailing membrane
  filtration processes that can be used to separate solutes on the order of 1 nm
  in size and smaller, including ions. However, commercially available amorphous
  RO membranes and porous NF membranes lack uniformity, which severely impacts
  their selectivity \cite{van_der_bruggen_review_2003}. Although scalable, their
  fabrication involves the spontaneous assembly of polymers into disordered
  structures that offer separation pathways that are tortuous and polydisperse in
  size \cite{werber_materials_2016}. Tortuosity increases the effective length
  that a solute must travel, while pore size polydispersity limits membrane
  selectivity.

@@

  Systems built with 5 columns around each pore equilibrate to a pore spacing
  that is most consistent with the experimental value of 4.12 nm derived from
  SAXS measurements (Figure~\ref{fig:SAXS}). Ordered basin systems built with 4
  columns per pore equilibrate to an average pore spacing 0.25 nm lower than
  experiment. Ordered basin systems built with 6 columns per pore, have an
  equilibrated pore spacing ca. 0.50 nm higher than experiment. Monomers in
  disordered basin systems built with 6 columns-per-pore agree with experimental
  pore-to-pore distances within error, but stack too far apart.  Monomers in 6
  column-per-pore disordered sandwiched and disordered parallel displaced
  configurations stack $\sim$~4.87 and 4.94 \AA~apart respectively, which is
  $\sim$ 1.2~\AA~further apart than suggested by experiment. 5 column-per-pore
  systems stack, at a maximum, 0.9 \AA~further apart than experiment (see
  Table~\ref{table:correlation_length}). The remainder of this discussion will
  therefore focus on the analysis of systems built with 5 columns per pore.

@@

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{intensity_vs_zsigma.pdf}
  \caption{}\label{fig:intensity_vs_zsigma}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{rpi_xsection_vs_zsigma.pdf}
  \caption{}\label{fig:rpi_xsection_vs_zsigma}
  \end{subfigure}
  \caption{We can decrease the maximum intensity of R-$\pi$ by increasing
  thermal noise in the $z$-direction; however, the $q_r$ cross-section profile does
  not change. (a) The intensity of R-$\pi$ drops precipitously as we increase
  $\sigma_z$. We can decrease the intensity of R-$\pi$, relative to the
  background intensity, relative to systems with simulation levels of disorder (blue)
  by a factor of 3 when we increase $\sigma_z$ by 12\% (green) and by a factor 
  of 15 when we increase $\sigma_z$ by 27\% (red). (b) Although the maximum intensity of R-$\pi$
  decreases with increasing $\sigma_z$, the peak shape remains the same.}\label{fig:znoise}
  \end{figure}

@@

  It therefore seems likely that the discrepancy between the shape and intensity of
  R-$\pi$ generated from our simulations versus experiment is primarily due to
  highly correlated columns in our highly symmetric initial configurations.
  If the columns were uncorrelated, $g(z)$ of the ordered sandwiched system 
  would very closely resemble Figure~\ref{fig:z_correlation_sandwich} when we 
  average all $z$-slices of the 3D correlation function. Contributions to
  $g(z)$ from head groups in other columns within the same pore, and with head
  groups in different pores would cancel out. However, when we include all 
  $z$-slices in $g(z)$ for our system, rather than restricting it to those 
  $z$-slices within 2.1~\AA of the center of the correlation function 
  (Section~\ref{section:correlation_length}), $g(z)$ shows oscillatory 
  behavior and its correlation length nearly triples to 12.4 $\pm$ 0.7 \AA.
  (Figure~\ref{fig:z_correlation_fullbox}). The peaks of $g(z)$ are actually 
  reinforced by head groups in neighboring columns which is not expected
  in a system where columns act independently.

  \begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{z_correlation_fullbox.pdf}
  \caption{When we average all $z$-slices of the 3D correlation function, the 
  correlation length of the ordered sandwich configuration nearly triples.}\label{fig:z_correlation_fullbox}
  \end{figure}

  Increasing thermal noise in the $xy$ plane causes the FWHM of the $q_r$
  cross-section of R-$\pi$ to, somewhat counterintuitively, decrease. Using a
  system with uncorrelated columns, we modified the disorder in both the $r$ and
  $\theta$ directions by multiplying their values by a factor of 0.5 and 2
  (Figure~\ref{fig:qy_fwhm}). When we cut the noise in half, the FWHM increases
  by 88\%. When we double the noise, the FWHM decreases by 51\%. As we have fit
  the data in Figure~\ref{fig:sim_rsection_fit}, the FWHM of R-$\pi$ agrees well
  with experiment, meaning the quenched configuration contains about the right
  amount of disorder on the $xy$ plane. However, the relatively low quality of
  the fit implies that the FWHM of a fixed Lorentzian fit to data from a single
  simulation might not be the best comparison to experiment and that some form of
  initial configuration dependence causes extra structure.

@@

  Since the dynamics of this system are slow, consequently most of our analysis
  has been focused on a set of quenched configurations with some structural
  characteristics that do not vary significantly with time. We therefore
  attempted to learn more about their time-averaged structures by studying
  ensembles of shorter simulations started from more independent initial
  configurations. To do this, we created ensembles of 40 simulations in both the
  parallel displaced and sandwiched configurations. We used the same
  equilibration scheme described in Section~\ref{section:equilibration}.
  However, we built the initial configurations with columns randomly displaced in
  the $z$-direction and looked at the results after 5 ns of unrestrained
  simulation, which should capture any quenched disorder built in from the
  initial configurations. We then simulated XRD patterns from the 40
  trajectories, concatenated together. The $q_r$ cross-sections of R-$\pi$
  generated from the two concatenated trajectories are shown in
  Figure~\ref{fig:ensemble_XRD}.

@@

  These short independent simulations exhibit a symmetric distribution of
  quenched disorder providing evidence that the quenched disorder is randomized
  in each system.  The only difference between each independent initial
  configuration is the random $z$-direction displacement that we apply to each
  column. The quenched disorder is then a result of the initial structural
  collapse that is directed by the randomized velocity vectors of the system's
  atoms drawn from a Maxwell-Boltzmann distribution. We pooled the deviations of
  each head group COM from idealized positions from all independent simulations
  in order to create the distributions shown in
  Figure~\ref{fig:offset_ensemble_pooled} (see
  Section~\ref{method:simple_systems}). The distributions appear symmetric
  meaning that there is an equal chance for a head group to displace in the
  positive or negative $z$, $r$, or $\theta$-direction.

  We verified that this same symmetry exists for individual COMs.  If we
  consider the COM of a given head group, we expect that the distribution of its
  deviations from its idealized position among all the independent simulations
  should be symmetric about its idealized position. We can calculate the
  empirical cumulative distribution function (ECDF) of these distributions of
  deviations. The ECDFs of 3 randomly selected COMs are shown in
  Figure~\ref{fig:offset_ecdfs} to guide the eye (colored lines). The average of
  all COM ECDFs is represented by the red line with the 95\% confidence interval
  represented by the red error bars. We also calculated the ECDF from the full
  distributions in Figure~\ref{fig:offset_ensemble_pooled} (black line with error
  bars). We generated error bars that represent the 95\% confidence intervals by
  performing 1000 trials where we randomly sampled 40 values (the same as the
  number of independent simulations) from the distributions in
  Figure~\ref{fig:offset_ensemble_pooled}, then measured the deviation between
  the sampled ECDF and the full ECDF. These error bars are generally smaller than
  those generated from the 400 COM ECDFs, as there is a wider distribution of
  mean COM deviations than the mean deviations from distributions sampled from
  the full distribution. The similarity between entirely independently sampled
  ECDFs and the ECDFs from individual independent simulations supports the
  hypothesis that each quenched configuration is random and mostly independent.

@@

@@

  We observe slow dynamics in our system. Typical diffusion constants for
  columnar liquid crystals have been reported to be between $10^{-11}~
  \mathrm{m}^2/\mathrm{s}$ \cite{dong_translational_1984} and
  $10^{-14}~\mathrm{m}^2/\mathrm{s}$ \cite{dvinskikh_molecular_2002}.  In the
  case of LLCs, we would expect a diffusion coefficient at the low end of this
  range since hydrophilic regions of monomers can hydrogen bond, thus increasing
  the energetic barriers for motion. We measured the diffusion constants of
  monomers in each of the main systems we studied (see Table \ref{S-table:msd} of
  the SI) and find they are all on the order of
  $10^{-14}~\mathrm{m}^2/\mathrm{s}$. However, since on this timescale, no
  monomers move a full `level' above or below, this may be an overestimate. This
  slow speed is not entirely surprising considering how densely the monomers are
  packed as implied by diffraction experiments. The entangled tails likely
  restrict both vertical and lateral movement.

@@

  \clearpage
  \bibliography{llc}

@@

  \section*{Graphical TOC Entry}
  \begin{figure}
  \includegraphics[width=4.5cm]{toc_image_final.pdf}
  \includegraphics[width=4.07cm]{toc_waxs.pdf}
 \end{figure}
